---
title: "Traditional ML Models"
author: "Jinchen Xie"
date: "07/27/2020"
output: 
  html_document:
    theme: cerulean
    highlight: tango
    toc: true
    toc_depth: 3
    number_sections: true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(knitr)
library(tidyverse)
library(Hmisc)
library(caret)
library(glmnet)
library(mltools)
library(data.table)
library(pROC)
# library(boot)
# library(scales)
# library(gridExtra)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Model setup

## Import Data
```{r Import Data}
baseline.1 <- readxl::read_excel("/Users/jinchenxie/Desktop/datasets/processed/All Baseline Attributes combined.xlsx")
gbtm.df <- readxl::read_excel('~/Desktop/datasets/processed/GBTM_data/combined 7 labs data.xlsx')

# df contains only the sub-cohort of patients satisfied the lab measurements criteria.
df <- baseline.1[which(baseline.1$rID %in% gbtm.df$patients_id),]

rm(baseline.1, gbtm.df)
```


## Data Preprocessing
We want to include the same set of attributes that are considered in the GBTM model prediction. 

```{r Generate lists of attributes}
attris.df <- df %>% select("rID", "AGE_ADMISSION", "FEMALE", "RACE_White", starts_with("TOBACCO_"), starts_with("INSUR_"), starts_with("BP_"), "PULSE", "BMI", ends_with("_HST"), starts_with("HX_"), starts_with("PRIOR_"),  "CCI_TOTAL_SCORE", starts_with("CCI_"), ends_with("_CARDIOMYOPATHY"), ends_with("_00"), "CTB", ends_with("_CTB"), total_adms, Corrected_Followup_Days)

# Remove attributes that no patients had it as 1
no.patient.attris <- c("PRIOR_STERNOTOMY", "TOBACCO_STATUS_LABEL", "VIRAL_CARDIOMYOPATHY", "TZD_00", "GLP1_00", "SULFONYLUREA_00")
for (var in no.patient.attris){
  attris.df[[var]] <- NULL
}
rm(no.patient.attris, var)

nu.list <- c("AGE_ADMISSION", "CCI_TOTAL_SCORE", "BP_SYSTOLIC", "BP_DIASTOLIC", "PULSE", "BMI")
labels.list <- colnames(attris.df)[(dim(attris.df)[2]-7):dim(attris.df)[2]]
binary.list <- colnames(attris.df)[-1]
binary.list <- binary.list[(!binary.list %in% nu.list) & (!binary.list %in% labels.list)]
attris.list <- c(nu.list, binary.list)

variables.sum <- vector(mode="list")
variables.sum[["binary"]] <- binary.list
variables.sum[["numeric"]] <- nu.list
variables.sum[["label"]] <- labels.list
           
rm(attris.df, binary.list, nu.list, labels.list)
```

- The set of continuous, numeric attributes includes: `r nu.list`  

- The set of binary attributes include: `r binary.list`  

- The set of outcomes includes: `r labels.list[1:6]`  


### Avoid using highly correlated variables

Correlations > 0.5:  
```{r generate correlations table}
temp.df <- df %>% select(attris.list)
rcor1 <- rcorr(as.matrix(temp.df))
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
corr.df <- flattenCorrMatrix(rcor1$r, rcor1$P)

corr.df.5 <- corr.df[which(corr.df$cor > 0.5),]
corr.df.7 <- corr.df[which(corr.df$cor > 0.7),]
# corr.list <- union(corr.df.7$row, corr.df.7$column)
corr.df.5
rm(rcor1, corr.df.5, corr.df.7, temp.df, flattenCorrMatrix, corr.df)
```

```{r correlation plots matrix, eval=FALSE }
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
corr.df <- ctb.df %>% select(corr.list)
pairs(~., data = corr.df, lower.panel = panel.cor)
rm(corr.list, corr.df, panel.cor)
```

Remove highly correlated variables CCI_CHF, CCI_COPD, CCI_DM_NO_CC  
Variables removed due to high correlations (>0.7)
```{r remove correlated variables}
# remove variables
rm.attris <- c("CCI_CHF", "CCI_COPD", "CCI_DM_NO_CC")
for (var in rm.attris){
  df[[var]] <- NULL
}
attris.list <- attris.list[!attris.list %in% rm.attris]
rm(rm.attris, var)
```


```{r Create oucomes data.frames, eval=FALSE }
# Generate 6 dataframes based on different variable as label
ctb.df <- df %>% select(attris.list, "CTB")
in.ctb.df <- df %>% select(attris.list, "in_CTB")
ctb.30.df <- df %>% select(attris.list, "30_day_CTB")
ctb.60.df <- df %>% select(attris.list, "60_day_CTB")
ctb.90.df <- df %>% select(attris.list, "90_day_CTB")
ctb.1yr.df <- df %>% select(attris.list, "1_year_CTB")
```

### Check class imbalance
```{r eval=FALSE, echo=TRUE}
table(ctb.df$CTB)/dim(df)[1]
table(in.ctb.df$in_CTB)/dim(df)[1]
table(ctb.30.df$`30_day_CTB`)/dim(df)[1]
table(ctb.60.df$`60_day_CTB`)/dim(df)[1]
table(ctb.90.df$`90_day_CTB`)/dim(df)[1]
table(ctb.1yr.df$`1_year_CTB`)/dim(df)[1]
```




## Data Splitting
Split data into train and test sets, for reproducible results, we set seed = 123

```{r train test split}
smp_size <-  floor(0.75*nrow(ctb.df))
set.seed(123)
train.ind <- sample(seq_len(nrow(ctb.df)), size = smp_size)

rm(smp_size)
```

There are `r nrow(train)` samples in training set, and `r nrow(test)` samples in test set. 




# Logistic Regression

## Full model

*All-cause CTB as label:  *
```{r}
train <- ctb.df[train.ind, ]
test <- ctb.df[-train.ind, ]
```

```{r eval=FALSE}
# for (i in seq(1, length(names(train))-1)){
  # print(i)
  # print(summary(glm(CTB ~ eval(as.name(names(train)[i])), data = train, family = binomial))$coefficient)
}  
```

**Predictors that are associated with CTB in bivariate analysis, we get the following variables have significant p-value (<0.2 cutoff).**  
AGE_ADMISSION, CCI_TOTAL_SCORE, BP_DIASTOLIC, FEMALE, RACE_White, TOBACCO_Never, `TOBACCO_Q/Remote`, INSUR_Commercial, INSUR_Medicaid, INSUR_Medicare, HTN_HST, HYPERLIPIDEMIA_HST, DIAB_HST, COPD_HST, CAD_HST, AFIB_HST, PH_HST, VASCULARDISEASE_HST, CKD_HST, ESRD_HST, PRIOR_SMVR, CCI_MI, CCI_PERIPHERAL_VASC, CCI_DM_WITH_CC, CCI_RENAL_DISEASE, ASA_00, NOAC_00, APT_00, ACE_ARB_00, AAD_CLASS_I_00, INSULIN_00, METFORMIN_00

```{r model 1, echo=TRUE}
ctb.glm.fit1 <- glm(CTB ~ AGE_ADMISSION + CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + `TOBACCO_Q/Remote` + INSUR_Commercial + INSUR_Medicaid + INSUR_Medicare + HTN_HST + HYPERLIPIDEMIA_HST + DIAB_HST + COPD_HST + CAD_HST + AFIB_HST + PH_HST + VASCULARDISEASE_HST + CKD_HST + ESRD_HST + PRIOR_SMVR + CCI_MI + CCI_PERIPHERAL_VASC + CCI_DM_WITH_CC + CCI_RENAL_DISEASE + ASA_00 + NOAC_00 + APT_00 + ACE_ARB_00 + AAD_CLASS_I_00 + INSULIN_00 + METFORMIN_00, data = train, family = binomial)
summary(ctb.glm.fit1)


ctb.glm.probs <- predict(ctb.glm.fit1, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```

```{r}
ctb.df <- ctb.df %>% select(AGE_ADMISSION, CCI_TOTAL_SCORE, BP_DIASTOLIC, FEMALE, RACE_White, TOBACCO_Never, `TOBACCO_Q/Remote`, INSUR_Commercial, INSUR_Medicaid, INSUR_Medicare, HTN_HST, HYPERLIPIDEMIA_HST, DIAB_HST, COPD_HST, CAD_HST, AFIB_HST, PH_HST, VASCULARDISEASE_HST, CKD_HST, ESRD_HST, PRIOR_SMVR, CCI_MI, CCI_PERIPHERAL_VASC, CCI_DM_WITH_CC, CCI_RENAL_DISEASE, ASA_00, NOAC_00, APT_00, ACE_ARB_00, AAD_CLASS_I_00, INSULIN_00, METFORMIN_00, CTB) 

train <- ctb.df[train.ind, ]
test <- ctb.df[-train.ind, ]
```


## LASSO regression
Standardize continuous variables to scale 0-1.  
Min-Max Scaling is used.  
```{r min-max scaling}
train.stand <- train
test.stand <- test
preproc <- preProcess(train[,c(1:6)], method=c("center", "scale"))
train.stand[,c(1:6)] <- predict(preproc, train[,c(1:6)])
preproc <- preProcess(test[,c(1:6)], method=c("center", "scale"))
test.stand[,c(1:6)] <- predict(preproc, test[,c(1:6)])
rm(preproc)
```

```{r}
train.x <- model.matrix(CTB~., data = train.stand)[,-1] 
train.y <- train.stand$CTB

glm.lasso <- glmnet(train.x, train.y)

plot(glm.lasso, xvar="lambda", label=TRUE, ylim=c(-0.1, 0.1))

rm(train.x, train.y)
```

For the lambda = `r glm.lasso$lambda[5]`: 
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[5])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[5])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[5])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[5])@i]`.  

For the lambda = `r glm.lasso$lambda[10]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[10])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[10])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[10])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[10])@i]`.  

For the lambda = `r glm.lasso$lambda[15]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[15])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[15])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[15])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[15])@i]`.  

For the lambda = `r glm.lasso$lambda[20]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[20])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[20])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[20])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[20])@i]`.

We use the attributes For the lambda = `r glm.lasso$lambda[20]`:  
```{r model 2, echo=FALSE}
ctb.glm.fit2 <- glm(CTB ~ AGE_ADMISSION + CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + `TOBACCO_Q/Remote` + INSUR_Commercial + DIAB_HST + COPD_HST + CAD_HST + PH_HST + ESRD_HST + PRIOR_SMVR + CCI_PERIPHERAL_VASC + CCI_DM_WITH_CC + CCI_RENAL_DISEASE + NOAC_00 + APT_00 + ACE_ARB_00 + INSULIN_00 + METFORMIN_00, data = train, family = binomial)
summary(ctb.glm.fit2)
  
ctb.glm.probs <- predict(ctb.glm.fit2, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)
       
# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```


**The misclassification rate is** `r ctb.test.mis*100`%  

```{r model 3, echo=TRUE}
ctb.glm.fit3 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never +  INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC +  NOAC_00 + INSULIN_00, data = train, family = binomial)
summary(ctb.glm.fit3)

ctb.glm.probs <- predict(ctb.glm.fit3, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```





*30-day CTB as label:  *
```{r}
train <- ctb.30.df[train.ind, ]
test <- ctb.30.df[-train.ind, ]
```

```{r}
ctb.glm.fit6 <- glm(`30_day_CTB` ~ ., data = train, family = binomial)
summary(ctb.glm.fit6)


ctb.glm.probs <- predict(ctb.glm.fit6, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`30_day_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

# The confusion matrix of the prediction is   
confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`30_day_CTB`))

roc_obj <- pROC::roc(test$`30_day_CTB`, ctb.glm.pred)
auc(roc_obj)
```

*in-hospital CTB as label:  *
```{r}
train <- in.ctb.df[train.ind, ]
test <- in.ctb.df[-train.ind, ]
```

```{r model 11}
ctb.glm.fit11 <- glm(in_CTB ~ ., data = train, family = binomial)
summary(ctb.glm.fit11)


ctb.glm.probs <- predict(ctb.glm.fit11, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`in_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

# The confusion matrix of the prediction is   
confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`in_CTB`))

roc_obj <- pROC::roc(test$in_CTB, ctb.glm.pred)
auc(roc_obj)
```


 30-day CTB
```{r}
train <- ctb.30.df[train.ind, ]
test <- ctb.30.df[-train.ind, ]
```

Standardize continuous variables to scale 0-1.  
Min-Max Scaling is used.  
```{r min-max scaling}
train.stand <- train
test.stand <- test
preproc <- preProcess(train[,c(1:6)], method=c("center", "scale"))
train.stand[,c(1:6)] <- predict(preproc, train[,c(1:6)])
preproc <- preProcess(test[,c(1:6)], method=c("center", "scale"))
test.stand[,c(1:6)] <- predict(preproc, test[,c(1:6)])
rm(preproc)
```

```{r}
train.x <- model.matrix(`30_day_CTB`~., data = train.stand)[,-1] 
train.y <- train.stand$`30_day_CTB`

glm.lasso <- glmnet(train.x, train.y)

plot(glm.lasso, xvar="lambda", label=TRUE, ylim=c(-0.2, 0.2))
```

For the lambda = `r glm.lasso$lambda[5]`: 
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[5])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[5])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[5])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[5])@i]`.  

For the lambda = `r glm.lasso$lambda[10]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[10])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[10])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[10])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[10])@i]`.  

For the lambda = `r glm.lasso$lambda[15]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[15])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[15])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[15])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[15])@i]`.  

For the lambda = `r glm.lasso$lambda[20]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[20])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[20])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[20])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[20])@i]`.

```{r}
rm(train.x, train.y)
```


We use the attributes For the lambda = `r glm.lasso$lambda[10]`:  
```{r model 7, echo=FALSE}
ctb.glm.fit7 <- glm(`30_day_CTB` ~ AGE_ADMISSION + CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + INSUR_Commercial + INSUR_Medicare + COPD_HST + CAD_HST + ESRD_HST + PRIOR_ENDO_AO_GRFT + PRIOR_STVR + CCI_PERIPHERAL_VASC + CCI_CEREBROVASCULAR + CCI_RENAL_DISEASE + BB_00 + ACE_ARB_00 +  METFORMIN_00, data = train, family = binomial)
summary(ctb.glm.fit7)

ctb.glm.probs <- predict(ctb.glm.fit7, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`30_day_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`30_day_CTB`))

roc_obj <- pROC::roc(test$`30_day_CTB`, ctb.glm.pred)
auc(roc_obj)
```



```{r model 8, echo=FALSE}
ctb.glm.fit8 <- glm(`30_day_CTB` ~ CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + INSUR_Commercial  + COPD_HST + ESRD_HST + CCI_PERIPHERAL_VASC + CCI_CEREBROVASCULAR  + BB_00 + ACE_ARB_00 +  METFORMIN_00, data = train, family = binomial)
summary(ctb.glm.fit8)

ctb.glm.probs <- predict(ctb.glm.fit8, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`30_day_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`30_day_CTB`))

roc_obj <- pROC::roc(test$`30_day_CTB`, ctb.glm.pred)
auc(roc_obj)
```

 in-hospital CTB

```{r}
train <- in.ctb.df[train.ind, ]
test <- in.ctb.df[-train.ind, ]
```

Standardize continuous variables to scale 0-1.  
Min-Max Scaling is used.  
```{r min-max scaling}
train.stand <- train
test.stand <- test
preproc <- preProcess(train[,c(1:6)], method=c("center", "scale"))
train.stand[,c(1:6)] <- predict(preproc, train[,c(1:6)])
preproc <- preProcess(test[,c(1:6)], method=c("center", "scale"))
test.stand[,c(1:6)] <- predict(preproc, test[,c(1:6)])
rm(preproc)
```

```{r}
train.x <- model.matrix(`in_CTB`~., data = train.stand)[,-1] 
train.y <- train.stand$`in_CTB`

glm.lasso <- glmnet(train.x, train.y)

plot(glm.lasso, xvar="lambda", label=TRUE, ylim=c(-0.1, 0.1))
```

For the lambda = `r glm.lasso$lambda[3]`: 
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[5])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[5])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[3])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[3])@i]`.  

For the lambda = `r glm.lasso$lambda[10]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[10])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[10])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[10])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[10])@i]`.  

For the lambda = `r glm.lasso$lambda[15]`:  
There are `r length((coef(glm.lasso, s = glm.lasso$lambda[15])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[15])@i])` variables that have non-zero coefficients in the model . Namely, they are `r (coef(glm.lasso, s = glm.lasso$lambda[15])@Dimnames[[1]][-1])[coef(glm.lasso, s = glm.lasso$lambda[15])@i]`.  

```{r}
rm(train.x, train.y)
```


We use the attributes For the lambda = `r glm.lasso$lambda[10]`:  
```{r model 12, echo=FALSE}
ctb.glm.fit12 <- glm(`in_CTB` ~ CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + CKD_HST + CAD_HST + CIRRHOSIS_HST+ PRIOR_ENDO_AO_GRFT + PRIOR_STVR + CCI_MI + ACE_ARB_00 , data = train, family = binomial)
summary(ctb.glm.fit12)

ctb.glm.probs <- predict(ctb.glm.fit12, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`in_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`in_CTB`))

roc_obj <- pROC::roc(test$`in_CTB`, ctb.glm.pred)
auc(roc_obj)
```


## Add Last lab reading
```{r}
ctb.labs.df <- cbind(ctb.df, ave.last.df)
train <- ctb.labs.df[train.ind, ]
test <- ctb.labs.df[-train.ind, ]
```

```{r}
ctb.glm.fit14 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC + NOAC_00 + INSULIN_00 + EGFR_last + N_A_last, data = train, family = binomial)
summary(ctb.glm.fit14)

ctb.glm.probs <- predict(ctb.glm.fit14, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```


## Add Group info
We add the group information got from GBTM multi variables model into the logistic regression (LASSO regulariation). We add the group number as a factor.  

### ALL-CAUSE CTB
```{r}
#  initialize ctb dataframe with group info
ctb.gr.df <- ctb.df
```

```{r}
ctb.gr.df <- readxl::read_excel('~/Desktop/datasets/processed/GBTM_data/ctb_w_groupinfo.xlsx', sheet="Sheet 1")
```


```{r}
# import gbtm results saved in STATA
lab.gbtm.df <- readxl::read_excel("~/Desktop/trial save.xlsx", sheet="Sheet1")

# add the group number as a variable in the ctb.df
ctb.gr.df <- cbind(ctb.gr.df, factor(lab.gbtm.df$`_traj_Group`))
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_hgb"
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_egfr"
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_wbc"
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_k"
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_na"
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_platelets"
# names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_glucose"
names(ctb.gr.df)[names(ctb.gr.df) == "factor(lab.gbtm.df$`_traj_Group`)"] <- "group_multivar"
```

group_multivar3: 4 labs multivariate model with HGB, EGFR, NA, WBC
group_multivar4: 5 labs multivariate model with HGB, EGFR, NA, K, WBC 
group_multivar5: 4 labs multivariate model without HGB, EGFR, NA, K 
group_multivar6: 3 labs multivatiate model with HGB, EGFR, NA
group_multivar7: 6 labs multivatiate model with all significant baseline covariates and time-varying covariates POST_VAD

```{r}
temp.df <- ctb.gr.df %>% select(starts_with("group_"))
rcor1 <- rcorr(as.matrix(temp.df))
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
corr.df <- flattenCorrMatrix(rcor1$r, rcor1$P)

corr.df.7 <- corr.df[which(corr.df$cor > 0.7),]
corr.df.7
rm(temp.df, rcor1, flattenCorrMatrix, corr.df, corr.df.7)
```



```{r}
xtabs(~ group_egfr + group_na, ctb.gr.df)
```




```{r}
ave.last.df <- readxl::read_excel('~/Desktop/datasets/processed/GBTM_data/last and average labs data.xlsx') %>% select(ends_with("_last"), ends_with("_average"), ends_with("_min"), ends_with("_max"), ends_with("_sd"))

combined.ctb.gr.df <- cbind(ctb.gr.df, ave.last.df)


train <- combined.ctb.gr.df[train.ind, ]
test <- combined.ctb.gr.df[-train.ind, ]
```

```{r}
ctb.gr.df[,which(names(ctb.gr.df) == "group_egfr_3")] <- NULL
```


```{r save the group info for each labs}
openxlsx::write.xlsx(combined.ctb.gr.df, "~/Desktop/datasets/processed/prediction_data/ctb_preprocessed.xlsx", col.names = TRUE, row.names = FALSE)


write.xlsx(ctb.gr.df, '~/Desktop/datasets/processed/GBTM_data/ctb_w_groupinfo.xlsx', col.names = TRUE, row.names = FALSE)

write.xlsx(combined.ctb.gr.df, '~/Desktop/datasets/processed/GBTM_data/ctb_w_groupinfo_lastave.xlsx', col.names = TRUE, row.names = FALSE)
```

```{r}
# ggplot(data = ctb.gr.df) + geom_bar(mapping = aes(fill = group_egfr_na, x = `CTB`))
```


```{r model 4, echo=TRUE}
ctb.glm.fit4 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC +  NOAC_00 + INSULIN_00 + group_egfr + group_na, data = train, family = binomial)
summary(ctb.glm.fit4)

ctb.glm.probs <- predict(ctb.glm.fit4, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))


roc_obj <- pROC::roc(response = test$CTB, predictor = ctb.glm.pred)
auc(roc_obj)
```

```{r model 5, echo=TRUE}
ctb.glm.fit5 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC + NOAC_00 + INSULIN_00 + group_egfr + group_na + EGFR_last + EGFR_average, data = train, family = binomial)
summary(ctb.glm.fit5)

ctb.glm.probs <- predict(ctb.glm.fit5, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))


roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```

```{r model 6, echo=TRUE }
ctb.glm.fit6 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC + NOAC_00 + INSULIN_00 + group_egfr + group_na + EGFR_last + EGFR_average, data = train, family = binomial)
summary(ctb.glm.fit6)

ctb.glm.probs <- predict(ctb.glm.fit6, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```

```{r model 7, echo=TRUE }
ctb.glm.fit7 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC + NOAC_00 + INSULIN_00 + group_egfr_na, data = train, family = binomial)
summary(ctb.glm.fit7)

ctb.glm.probs <- predict(ctb.glm.fit7, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```

```{r}
ctb.glm.fit6 <- glm(CTB ~ AGE_ADMISSION + BP_DIASTOLIC + FEMALE + RACE_White + TOBACCO_Never + INSUR_Commercial + COPD_HST + CCI_PERIPHERAL_VASC + NOAC_00 + INSULIN_00 + group_egfr + group_na, data = train, family = binomial)
summary(ctb.glm.fit6)

ctb.glm.probs <- predict(ctb.glm.fit6, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$CTB)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$CTB))

roc_obj <- pROC::roc(test$CTB, ctb.glm.pred)
auc(roc_obj)
```


### 30-day CTB

```{r}
ctb.30.gr.df <- cbind(ctb.gr.df, ctb.30.df$`30_day_CTB`)
names(ctb.30.gr.df)[names(ctb.30.gr.df) == "ctb.30.df$`30_day_CTB`"] <- "30_day_CTB"
```

```{r}
train <- ctb.30.gr.df[train.ind, ]
test <- ctb.30.gr.df[-train.ind, ]
```

```{r model 9, echo=TRUE}
ctb.glm.fit9 <- glm(`30_day_CTB` ~ CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + INSUR_Commercial  + COPD_HST + ESRD_HST + CCI_PERIPHERAL_VASC + CCI_CEREBROVASCULAR  + BB_00 + ACE_ARB_00 +  METFORMIN_00 + group_wbc + group_k + group_na + group_hgb + group_egfr + group_platelets, data = train, family = binomial)
summary(ctb.glm.fit9)

ctb.glm.probs <- predict(ctb.glm.fit9, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`30_day_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`30_day_CTB`))

roc_obj <- pROC::roc(test$`30_day_CTB`, ctb.glm.pred)
auc(roc_obj)
```


### in-hospital CTB
```{r}
ctb.in.gr.df <- cbind(ctb.gr.df, in.ctb.df$in_CTB)
names(ctb.in.gr.df)[names(ctb.in.gr.df) == "in.ctb.df$in_CTB"] <- "in_CTB"
```

```{r}
train <- ctb.in.gr.df[train.ind, ]
test <- ctb.in.gr.df[-train.ind, ]
```

```{r model 13, echo=TRUE}
ctb.glm.fit13 <- glm(`in_CTB` ~ CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + CKD_HST + CAD_HST + CIRRHOSIS_HST+ PRIOR_ENDO_AO_GRFT + PRIOR_STVR + CCI_MI + ACE_ARB_00 + group_wbc + group_hgb + group_k + group_na + group_platelets, data = train, family = binomial)
summary(ctb.glm.fit13)

ctb.glm.probs <- predict(ctb.glm.fit13, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`in_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`in_CTB`))

roc_obj <- pROC::roc(test$`in_CTB`, ctb.glm.pred)
auc(roc_obj)
```

```{r model 14, echo=TRUE}
ctb.glm.fit14 <- glm(`in_CTB` ~ CCI_TOTAL_SCORE + BP_DIASTOLIC + FEMALE + CKD_HST + CAD_HST + CIRRHOSIS_HST+ PRIOR_ENDO_AO_GRFT + PRIOR_STVR + CCI_MI + ACE_ARB_00 + group_multivar, data = train, family = binomial)
summary(ctb.glm.fit14)

ctb.glm.probs <- predict(ctb.glm.fit14, test, type="response")
ctb.glm.pred <- rep(0, length(ctb.glm.probs))
ctb.glm.pred[ctb.glm.probs > 0.5] <- 1
tab.glm <- table(ctb.glm.pred, test$`in_CTB`)

# Misclassification rate
ctb.test.mis <- sum(tab.glm[row(tab.glm)!=col(tab.glm)])/nrow(test)

confusionMatrix(data = factor(ctb.glm.pred), reference = factor(test$`in_CTB`))

roc_obj <- pROC::roc(test$`in_CTB`, ctb.glm.pred)
auc(roc_obj)
```


# Distal outcome model

```{r}
gbtm.outcome.df <- readxl::read_excel("~/Desktop/trial save.xlsx", sheet="Sheet1")
gbtm.outcome <- gbtm.outcome.df[-train.ind, ]$`_traj_Outcome`
gbtm.pred <- rep(0, length(gbtm.outcome))
gbtm.pred[gbtm.outcome > 0.5] <- 1
tab.gbtm <- table(gbtm.pred, test$`CTB`)
confusionMatrix(data = factor(gbtm.pred), reference = factor(test$CTB))
roc_obj <- pROC::roc(test$CTB, gbtm.pred)
auc(roc_obj)
```

